### 2019-06-04 code/select_usa_aeronet.Rmd
from Nov. 2018

---
title: "Select Aeronet in USA"
author: "Johnathan Rush"
date: "11/1/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(results = 'hold')
```

```{r message=FALSE}
library(data.table)
library(rgdal)
library(sp)
library(rgeos)
library(raster)
library(rnaturalearth)
```

# Open Aeronet

This is using the updated AERONET sites.  
*Change the `aeronet_dir` to where you have the AERONET data downloaded to*

```{r aeronet2017}
aeronet_dir = "~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/aeronet/"
aeronetsites <- fread(paste0(aeronet_dir, "aeronet_locations_v3.txt"))
# rename columns to shorter names
setnames(aeronetsites, c("Site_Name", "lon", "lat", "elevm"))
rmarkdown::paged_table(aeronetsites)
```

Make the AERONET data a Spatial Points object
```{r aeronet_spatial}
aeronetData = aeronetsites[, .(Site_Name, elevm)]
#aeronetPoints = SpatialPoints(cbind(aeronetsites$lon, aeronetsites$lat))
aeronetPoints = SpatialPoints(aeronetsites[, .(lon, lat)])
aeronetSP = SpatialPointsDataFrame(aeronetPoints, aeronetData)
aeronetSP
```

Notice that we haven't set a Coordinate Reference System for the AERONET sites, yet.  
They're probably using WGS84 (the standard for GPS) datum lat/longs, so we'll set the CRS accordingly:

```{r aeronet_crs}
proj4string(aeronetSP) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
```

# Get a shape for the USA and for CONUS

First, download the Natural Earth data. This only has to be run once.  
*Change the `naturalEarthDir` to where you'd like to download this data.*

```{r set_ne_dir}
naturalEarthDir = "/data-belle/naturalearth/"
```
```{r dl_NE, eval=FALSE}
ne_download(scale = 10, type = "countries", category = "cultural", destdir = naturalEarthDir, load = FALSE)
```

Then load the countries file in as a Spatial Polygons object.  
```{r load_countries}
countriesSP = ne_load(scale = 10, type = "countries", category = "cultural", destdir = naturalEarthDir)
```

Plot all the aeronet stations worldwide
```{r plot_worldwide}
plot(countriesSP)
plot(aeronetSP, col = "red", add = TRUE)
```

Subset the countries to just the USA
```{r subset_usa}
usaSP = countriesSP[countriesSP$NAME == 'United States of America', "NAME"]
rm(countriesSP)
plot(usaSP)
```

Clip the USA down to just the CONUS - removing Alaska, Hawaii, PR, etc.  
```{r subset_conus}
clipmatrix = as.matrix(extent(usaSP))      # quick way to get a matrix with the right row/col names and dimensions
clipmatrix[] <- c(-126, 24.4, -66.7, 49.4) # change to actual coordinates of the CONUS bbox
clip_poly <- as(extent(as.vector(t(clipmatrix))), "SpatialPolygons") # CRS isn't set, but that's OK in this case
conusSP = gIntersection(usaSP, clip_poly, byid = TRUE)               # ignore the warning about different CRS
plot(conusSP)
```

# Select the AERONET stations in the USA

```{r aeronet_usa}
aeronet_USA = aeronetSP[usaSP, ]
aeronet_USA
plot(usaSP)
plot(aeronet_USA, col = "red", add = TRUE)
```

# Select the AERONET stations in the CONUS
```{r aeronet_conus}
aeronet_conus = aeronetSP[conusSP, ]
aeronet_conus
plot(conusSP)
plot(aeronet_conus, col = "red", add = TRUE)
```

# Export the AERONET selections

*All of these are set to export to the current working directory*. Change the `output_path` as needed.

After this are three different export verions. You may not need all of them.

```{r output_path}
output_path = getwd()
```

```{r hidden_path, echo=FALSE}
# this is just for the purpose of rendering the file to send to Meytar
output_path = "~/test"
```

First by ESRI Shapefile:  
```{r export_shapefile}
writeOGR(aeronet_USA, layer = "aeronet_usa", dsn = output_path, driver="ESRI Shapefile", overwrite_layer = TRUE)
writeOGR(aeronet_conus, layer = "aeronet_conus", dsn = output_path, driver="ESRI Shapefile", overwrite_layer = TRUE)
```

Then by Rds files for the SP objects:  
```{r export_rds}
saveRDS(aeronet_USA, file = paste0(output_path, "/aeronet_usa.Rds"))
saveRDS(aeronet_conus, file = paste0(output_path, "/aeronet_conus.Rds"))
```

Finally, by CSV files with the coordinates in the table:  
```{r export_csv}
aeronet_usa_DT = as.data.table(aeronet_USA)
aeronet_conus_DT = as.data.table(aeronet_conus)
fwrite(aeronet_usa_DT, file = paste0(output_path, "/aeronet_usa.csv"))
fwrite(aeronet_conus_DT, file = paste0(output_path, "/aeronet_conus.csv"))
```