### 2019-07-25 Code/04_AOD_join_MAIAC.Rmd

---
title: "AOD collocated with MAIAC"
author: "Yang Liu"
date: "07/16/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 7, fig.height = 4.5)
```

# Derive spatial variables within buffers
  st_buffer() to create circles around the aeronet stations (radius of 300km)
  Keep track of the AERONET ID, long/lat of station
  Intersect with base grid; but note that areas off the grid (e.g. ocean) can lead to very different numbers of cells within a buffer
  Long data.table with all pairLSTid within buffer per aeronet station
  Calculate distance from each grid centroid back to station
  Calculate variables at four buffer distances of 10km, 30km, 90km, and 270km
  - % non-missing
  -   Mean AOD
  -   AOD - Mean AOD

```{r}
source(here::here("Code", "funcs_for_plots.R"))
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
```

```{r}
# Terra sites in 2018
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra
site_sub <- unique(aod_lst_terra$Site_Name)
aod_site_NE <- readRDS(here("Intermediate", "aod_site_NE"))
# reproject to US atlas projection, so now the unit is meter 
aod_site_km <- st_transform(aod_site_NE[aod_site_NE$Site_Name%in%site_sub,], crs = 2163) 
st_crs(aod_site_km)

# Define radius (meter): 
radius0 = 300000
reference_lst_fst ="/data-belle/LST/MODIS.LST.C6/derived/conus_grid_201906.fst"
refDT = read_fst(reference_lst_fst, as.data.table = TRUE)[inNEMIA ==TRUE]
refDT_sf <- st_as_sf(refDT, coords = c("x_wgs84", "y_wgs84"), crs = 4326)
refDT_sf <-  st_transform(refDT_sf, crs = 2163)

# take each row (sf, df, geometry point) as the argument, create multi-point buffer 
# `aod_buffer` is a multi-buffer:
aod_buffers <- st_buffer(aod_site_km, radius0)
# mapview(aod_buffers) # if you plot this, you will see lots of circles (polygons)
# but of course, after intersect with the base grid, they won't be circles anymore. 
```

## Method

Note 7/22, for sure not the best way here, but it can work. 
1. Get geometry for buffers as a list: aod_buffers_list  
2. `st_intersects` buffers to reference grid, get a binary geometry list: join_list
3. get list of dataset of grid within each buffer: ref_list
4. Get geometry for points (site location) as a list: site_list2
5. Within each ref_list (a pie of points), Use `mapply` or a for loop to calculate distance from each grid point to the center (station)

@output: refDT_sub, AERONET sites matched to all the grid ids around it within the 300KM round buffer

```{r}
# 1. 
aod_buffers_list <- st_geometry(aod_buffers)
# 2. 
# join_list is a sparse geometry binary predicate lists, made of (empty) or 0L https://r-spatial.github.io/sf/reference/sgbp.html
join_list <- lapply(aod_buffers_list, function(x)st_intersects(refDT_sf, x))
# 3. 
# list of joined I (ID) for each Aeronet station in the grid `refDT_sf`
join_list_I <- lapply(join_list, function(x) sapply(x, function(z) if (length(z)==0) NA_integer_ else z[1]))
# ref_list is the grid points within each circle (NA removed)
ref_list <- lapply(join_list_I, function(x) refDT_sf[!is.na(x),])
# dist_list <- st_distance(ref_list, site_list[1:2])
# 4.   
site_list <- st_geometry(aod_site_km)  # for calculating distance 
# get level 1 list [i], level2 [[i]] is just a point pair
site_list2 <- lapply(1 : length(site_list), function(i) (site_list[i])) 
name_list <- as.list(aod_site_km$Site_Name) # for recording site names 
# 5. 
# select ref.grid subset by Index for each station
select.refDT.subset <- function(ref_list0, point0, name0){
  # ref_list0 = ref_list[[1]]
  # point0 = site_list2[[1]]; name0 = name_list[[1]]
  # cat(paste0("crs of ref grid is ", st_crs(ref_list0)$epsg,"; ",
  #            "crs of point is ", st_crs(point0)$epsg,"\n"))
  ref_list0$dist <- st_distance(ref_list0, point0)
  ref_list0$Site_Name <- name0
  return(as.data.table(ref_list0))
}

# # not sure why mapply doesn't work 
# refDT_sub_list <- mapply(select.refDT.subset, ref_list0 = ref_list[1], point0 = site_list2[1], name0 = name_list[1])

refDT_sub_list <- list()
for (i in 1:length(ref_list)){
  ref_list0 = ref_list[[i]];  point0 = site_list2[[i]]; name0 = name_list[[i]]
  ref_list0$dist <- st_distance(ref_list0, point0)
  ref_list0$Site_Name <- name0
  refDT_sub_list[[i]] <- ref_list0
}
refDT_sub <- rbindlist(refDT_sub_list)
refDT_sub[, dist_km:=as.numeric(dist)/1000] # change distance to km 
refDT_sub <- setDT(refDT_sub)[,.(Site_Name, idLSTpair0, dist_km)]
setnames(refDT_sub, "idLSTpair0", "idLSTpair0_in_buffer")
setkey(refDT_sub, Site_Name)
# save the dataset that links station to all the idLST in the round buffer
# refDT_sub: 3545991 x 3
saveRDS(refDT_sub, "~/Intermediate/AOD_data/190724_refDT_Site_idinBuffer.rds")

# view a circle
refDT_Wallop <- refDT_sub_list[[2]]
mapview::mapview(refDT_Wallop[sample(1:nrow(refDT_Wallop), nrow(refDT_Wallop)/100),])

```

## Calculation

@output new aod_lst_terra and aod_lst_aqua 
The collocated Aeronet with MODIS sites, plus the new variables

```{r}
refDT_sub <- readRDS("~/Intermediate/AOD_data/190724_refDT_Site_idinBuffer.rds")
# load the aeronet join MCD19
aod_lst_terra <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_terra
# the original MCD19 
lst_terra <- read.lst(sat = "terra", columns0 = c("idLSTpair0", "overpass_time", "Optical_Depth_047"))

# remove some variables 
aod.MODIS.newVars <- function(aod_join_MODIS, MODIS_all){
  aod_join_MODIS <- aod_join_MODIS[!is.na(Optical_Depth_047),]
  aod_join_MODIS[, day:=format(stn_time, "%Y-%m-%d")]
  var_time <- grep("time", names(aod_join_MODIS), value = T)
  aod_join_MODIS <- aod_join_MODIS[, -..var_time]

  setkey(aod_join_MODIS, Site_Name)
  aod_join_buffer <- aod_join_MODIS[refDT_sub, allow.cartesian = T]
  cat("The cartesian joined dataset between collocated Aeronet with MODIS sites data and base grid in each circles, dim is: ", paste(dim(aod_join_buffer),collapse = " x "))

  # join back to the MODIS file 
  MODIS_all[, day:=format(join_time, "%Y-%m-%d")]
  setnames(MODIS_all, c("idLSTpair0", "Optical_Depth_047"), c("idLSTpair0_in_buffer", "AOD_470_buffer"),
           skip_absent = T)
  setkey(MODIS_all, idLSTpair0_in_buffer, day)
  setkey(aod_join_buffer, idLSTpair0_in_buffer, day)
  # aod_join_buffer2 get AOD for all the buffer grid cells
  aod_join_buffer2 <- MODIS_all[,.(idLSTpair0_in_buffer, day, AOD_470_buffer)][aod_join_buffer]

  dist0 <- c(10, 30, 90, 270)
    calculate.vars.in.buffer <- function(distx){
      aod_join_buffer_new <- aod_join_buffer2[dist_km < distx, .(nonmissing = mean(!is.na(AOD_470_buffer)), AOD_buffer_mean = mean(AOD_470_buffer, na.rm = T)),
                              by =.(idLSTpair0, day)]
      setnames(aod_join_buffer_new, c("nonmissing", "AOD_buffer_mean"), 
               paste0(c("pNonNAAOD", "Mean_AOD"),distx, "km"))
      setkey(aod_join_buffer_new, idLSTpair0, day)
      return(aod_join_buffer_new)
    }
  new_varlist <- invisible(lapply(dist0, calculate.vars.in.buffer))
  new_vars <- Reduce(merge, new_varlist)
  setkey(new_vars, idLSTpair0, day)
  setkey(aod_join_MODIS, idLSTpair0, day)
  aod_join_MODIS <- new_vars[aod_join_MODIS]
  return(aod_join_MODIS)
}
aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
head(aod_lst_terra)
# AOD - AOD mean, do later 
saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")

aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
aod_lst_aqua <- aod.MODIS.newVars(aod_lst_aqua, lst_aqua)
head(aod_lst_aqua)
saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")

```

```{r}
# some summary
library(ggplot2)
plot.scatter.diagonal(aod_lst_terra, "Mean_AOD10km", "AOD_470nm")
cor(as.matrix(aod_lst_terra[,c(grep("Mean_AOD", names(aod_lst_terra), value = T), "AOD_470nm"), with = F]))
pmissing <- grep("pNon", names(aod_lst_terra), value = T)
summary(aod_lst_terra[,..pmissing])
```

```{r, eval = F}
# backup
# view the circle (buffer)
library(mapview)
mapview(st_buffer(aod_site_km[2,], radius0))

# distance is calculated differently under different project
st_distance(refDT_sf3[10,], aod_site_km[1,])
st_distance(st_transform(refDT_sf3[10,], crs= 4326), st_transform(aod_site_km[1,], crs= 4326))
geosphere::distm(as_Spatial(st_transform(refDT_sf3[10,], crs= 4326)), as_Spatial(aod_site_NE[1,]))

# An alternative way to join:
# refDT_sp <- as_Spatial(refDT_sf2)
# buf_poly <- as_Spatial(aod_buffer)
# sp_grid <- over(refDT_sp, buf_poly)

```