### 2019-07-29 Code/03.2_AOD Variation.Rmd
AERONET daily error

---
title: "AODV3 Level2 AOD2.0 from Aeronet: Day Variation"
author: "Yang Liu"
date: "7/24/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 8, fig.height = 4)
```

```{r}
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
source(here::here("Code", "funcs_for_plots.R"))
# compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit#

```

# Comparing within day variation of AERONET sites AOD measurements

@output dataset has dailyN: observations with in a day (at least 20)
@d_mean and sd_rmse are daily mean and sd

```{r}
# US site list
aeronet_conus_dt <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/aeronet_conus_dt.rds")
aod20_data_pred470 <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAOD_470nm_190618.fst", as.data.table = T)

# year > 2000
aod_conus <- aod20_data_pred470[year(date)>=2000][site%in%aeronet_conus_dt$Site_Name][!is.na(AOD_470nm)]
# aod_conus <- aod_conus[year(date)>2000]
aod_conus[,date:=as.Date(date, "%Y-%m-%d")]
aod_conus[,dailyN:=.N, by = .(site, date)]
aod_conus <- aod_conus[dailyN>20,]

aod_conus[, month := factor(months(date, abbr = T),
                            levels = month.abb)]
  aod_conus[month%in%c("Dec","Jan","Feb"), season:="Winter"]
  aod_conus[month%in%c("Mar","Apr","May"), season:="Spring"]
  aod_conus[month%in%c("Jun","Jul","Aug"), season:="Summer"]
  aod_conus[month%in%c("Sep","Oct","Nov"), season:="Fall"]
  aod_conus[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]

# summary1: aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[,.(lat, lon, dailyN, season, mean_aod = mean(AOD_470nm), 
                                           sd_rmse = sd(AOD_470nm)), by = .(site, date)])

```

## daily SD AOD generally follows daily mean AOD, more outliers in the summer

Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date.  
Plot by time gives this figure:

```{r}
# show daily trend: AOD vs date, all data
ggplot(data = aod1, aes(x = date, y = sd_rmse, color = site)) +
  geom_point(size = 1) + 
  theme_classic() + theme(legend.position = "none") + 
  labs(y = "Daily SD(AOD) by station", x = "Date") + 
  ggtitle("Daily variation by station across time", subtitle = "(# Daily obs >20)")

```

Trend within day

```{r, fig.width = 12, fig.height = 10}
# limit to long stations
site1 <- aod_conus[,.N, by = site][N>3330, site]

d1 <- aod_conus[site%in%site1,.(lat, lon, hour, dailyN, season, AOD_470nm, 
                                MeanAOD = mean(AOD_470nm),
                                sd_rmse = sd(AOD_470nm)), by = .(site, date)]
d1[, stdAOD:=AOD_470nm - MeanAOD]
d1[, hour := as.numeric(hour)]
d1[hour%in%c(0,1,2), hour :=hour+24]
d1[, totalN := sum(dailyN), by = site]
setorder(d1, -totalN)
# data ready 
site2 <- aod_conus[,.N, by = site][N>8E4, site]
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = stdAOD, color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam") + 
  ggtitle("Daily variation (AOD - daily Mean AOD) by hour")

# sd(AOD) by hour, mean follows same trend
d1_hour <- d1[site%in%site2,.(sd_hour = sd(AOD_470nm), mean_hour = mean(AOD_470nm)), by = hour]
setkey(d1_hour, hour)
plot(x = d1_hour$hour, y = d1_hour$sd_hour, type = "o", ylab = "sd(AOD)")

#
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour")

#
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour")  + 
   facet_wrap(~season, ncol = 2)

# all sites

ggplot(dilute.data(d1,10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour") + 
  theme(legend.position = "none") + 
  facet_wrap(~season, ncol = 2)
```

## SD(AOD) by station

Then we calculated the median daily SD (very similar to mean) across all the days for each station.  
*N* adds up the number of all the hour-station observations _N = sum(dailyN)_  
mediansd = median(sd_rmse), meansd = mean(sd_rmse)

```{r}
aod2 <- unique(aod1[,.(mediansd = median(sd_rmse), N = sum(dailyN), lon, lat),
             by = .(site)])

setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2[N>5000 & mediansd > 0.02,])

N_label = 5E4
ggplot(data = aod2[N>3330,], aes(x = N, y = mediansd))+
    geom_point(size = 1, alpha = 1) + 
    theme_classic() + 
    ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = N, y = mediansd, label = site)) + 
    ggtitle("SD vs. Total Obs by Station", subtitle = ("Half the stations *")) + 
    labs(x = "Number of Observations (2000 - 2019)", y = "SD(AOD)")

```

* I have removed half of the stations with low total observations, that removed only `r  round(aod2[N<3330,sum(N)]/aod2[,sum(N)])` % of the AERONET AOD measurements (hour-station observation).

Generally speaking, smaller sites have higher variation, but still there are also some long-standing sites that have higher variation than others.

```{r, fig.width = 8, fig.height = 5}

ggplot() +
        geom_polygon(data = map_data("state"), 
                     aes(x = long, y = lat, group = group), 
                     fill = 'NA', color = "grey") + 
        geom_point(data = aod2[N>3330,],
                   aes(x=lon, y=lat, color = mediansd, size = N),
                   alpha = 0.8) + 
        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
        scale_size_area(max_size = 12) + 
        scale_colour_viridis_c() + 
        labs(x = "Longitude", y = "Latitude", color= "RMSE", 
             size = "Number of\nObservations") +
        ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = lon, y = lat, label = site)) + 
        theme_minimal() +
        ggtitle("SD by AERONET stations") +
        theme(legend.position ='right')

```

## Scatterplot and map By season

```{r, fig.width = 12, fig.height = 10}
aod1[, month := factor(months(date, abbr = T),
                            levels = month.abb)]
  aod1[month%in%c("Dec","Jan","Feb"), season:="Winter"]
  aod1[month%in%c("Mar","Apr","May"), season:="Spring"]
  aod1[month%in%c("Jun","Jul","Aug"), season:="Summer"]
  aod1[month%in%c("Sep","Oct","Nov"), season:="Fall"]
  aod1[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]

aod3 <- unique(aod1[,.(mediansd = median(sd_rmse), meansd = mean(sd_rmse), N = sum(dailyN), lon, lat),
             by = .(site, season)])

ggplot(data = aod3[N>3330,], aes(x = N, y = mediansd))+
    geom_point(size = 1, alpha = 1) + 
    theme_classic() + 
    facet_wrap(~ season, ncol = 2) + 
    ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = N, y = mediansd, label = site)) + 
    ggtitle("SD vs. Total Obs by Station", 
            subtitle = ("Showing long-lasting stations, half stations, 97% of the observations")) + 
    labs(x = "Number of Observations (2000 - 2019)", y = "SD by station")

ggplot() +
        geom_polygon(data = map_data("state"), 
                     aes(x = long, y = lat, group = group), 
                     fill = 'NA', color = "grey") + 
        geom_point(data = aod3[N>3330,],
                   aes(x=lon, y=lat, color = mediansd, size = N),
                   alpha = 0.8) + 
        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
        scale_size_area(max_size = 8) + 
        scale_colour_viridis_c() + 
        facet_wrap(~ season, ncol = 2) + 
        labs(x = "Longitude", y = "Latitude", color= "SD(AOD)", 
             size = "Number of\nObservations") +
        ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = lon, y = lat, label = site)) + 
        theme_minimal() +
        ggtitle("SD by AERONET stations") +
        theme(legend.position ='right')

```

## adjust for spline trend

```{r, eval=F}
library("splines")
site_selected <- aod_conus[,.N, by = site][N>3330, site]
aod3 <- aod_conus[site == site_selected[1]]
aod3[, dayint:=as.integer(date)]
aod3[, gam(AOD_470nm ~ s(dayint))]
fit0 <- smooth.spline(x = aod3$dayint, y = aod3$AOD_470nm)

plot(aod3[,.(dayint, AOD_470nm)], cex = 0.1)
lines(fit0, col = 'red')

```