### 2019-08-13 R/non_drake_code_and_md_files/03.2_AOD Variation.Rmd
organize folder

---
title: "AODV3 Level2 AOD2.0 Aeronet: investigate within-day Variation"
author: "Yang Liu"
date: "7/24/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 8, fig.height = 4)
```

```{r}
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
source(here::here("Code", "funcs_for_plots.R"))

```

compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit

# Comparing within day variation of AERONET sites AOD measurements

@output dataset has dailyN: observations with in a day (at least 20)
@sd_mean and sd_rmse are daily mean and sd

```{r}
# US site list
aeronet_conus_dt <- readRDS("~/Jdrive/PM/Just_Lab/projects/ECHO_PM/data/intermediate/aeronet_conus_dt.rds")
aod20_data_pred470 <- read.fst("/data-coco/ECHO_PM/AeronetAODV3Level2/AOD_Level20_All_Points_V3_wPredAOD_470nm_190618.fst", as.data.table = T)

# year > 2000
aod_conus <- aod20_data_pred470[year(date)>=2000][site%in%aeronet_conus_dt$Site_Name][!is.na(AOD_470nm)]
aod_conus[, year := year(date)]
# move the hours 0, 1, 2, attach the earlier date
aod_conus[, hour := as.numeric(hour)]
aod_conus[hour%in%c(0,1,2), `:=`(hour =hour+24, date = date - 1)]

# aod_conus <- aod_conus[year(date)>2000]
aod_conus[,datetime:=date]
aod_conus[,date:=as.Date(date, "%Y-%m-%d")]
aod_conus[,dailyN:=.N, by = .(site, date)] # obs in a day by site,day
aod_conus <- aod_conus[dailyN>=20,]
summary(aod_conus[,.N,by = .(site, date)][,N])

aod_conus[, month := factor(months(date, abbr = T),
                            levels = month.abb)]
  aod_conus[month%in%c("Dec","Jan","Feb"), season:="Winter"]
  aod_conus[month%in%c("Mar","Apr","May"), season:="Spring"]
  aod_conus[month%in%c("Jun","Jul","Aug"), season:="Summer"]
  aod_conus[month%in%c("Sep","Oct","Nov"), season:="Fall"]
  aod_conus[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]

```

* July has the most observations. Ndays are similar, what hours are shorter in winter

```{r}
# do winter have fewer observations per hour? 
aod_conus_hour <- aod_conus[,.(Nobs = .N, Ndays = uniqueN(date)), by = .(season, hour)]  
g1 = ggplot(aod_conus_hour, aes(x = hour, y = Ndays, color = season)) + geom_line() + theme_classic()
g2 = ggplot(aod_conus_hour, aes(x = hour, y = Nobs, color = season)) + geom_line() + theme_classic()
gridExtra::grid.arrange(grobs = list(g1, g2), ncol = 2, top = "By hour: similar days, fewer obs if daytime is short")

```

## Median Daily SD(AOD)

### calculating using all data

Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date. Then take the median.  @mediansd = median(sd_rmse)

```{r,  fig.width=12, fig.height=5}
# summary1: aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[,.(lat, lon, dailyN, month, season, 
                            mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
                            year = year(date)), by = .(site, date)])
aod1[,daysN:= uniqueN(date), by = site] # lasting days by site 
# 290 sites: aod2 [290 * 6]
aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site)])[yearmax>=2018]

# by season

aod3 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site, season)])[yearmax>=2018]

```

Among 290 stations, 87 have data from 2018 and 2019 (Figure Left).   
@daysN: days of the station with data

* The table shows daysN > 1500 stations.  
* `r aod2[daysN>=500,.N]` with `daysN` > 500 days are plotted  
* `r aod2[daysN>=1000,.N]` with `daysN` > 1000 days and SD(AOD)>0.013(the median)   are labelled in the map. Some other large stations are also labelled in the scatterplot.  
* So in the map, we label the top half of the long-lasting stations with at least 1000 days, and still on at least in 2018.

```{r, fig.width=12, fig.height=5}
# print results with daysN > 500 
setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2[daysN>1500])
library(ggrepel)
plotAODscatter <- function(aod_data, bySeason = F, daysN0 = 500){
  g1  <-  ggplot(data = aod_data[daysN>=daysN0], aes(x = daysN, y = mediansd))+
      geom_point(size = 1, alpha = 1) + 
      theme_classic() + 
      # ggrepel::geom_label_repel(data = aod_data[(daysN>daysN0*1.5 & mediansd> median(aod_data$mediansd))|daysN>daysN0*1.5], aes(x = daysN, y = mediansd, label = site)) + 
      ggtitle("Median daily SD(AOD) vs. Total days by Station", subtitle = ("Stations with latest year >=2018")) + 
      labs(x = "Number of Observations", y = "Daily SD(AOD)")
  if(bySeason) {g1 <- g1 + facet_wrap(~season)}
  g1
}

plotAODmap <- function(aod_data, bySeason = F, daysN0 = 500){
  g2 <- ggplot() +
          geom_polygon(data = map_data("state"), 
                       aes(x = long, y = lat, group = group), 
                       fill = 'NA', color = "grey") + 
          geom_point(data = aod_data[daysN >= daysN0],
                     aes(x=lon, y=lat, color = mediansd, size = daysN),
                     alpha = 0.8) + 
          # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
          scale_size_area(max_size = 10) + 
          scale_colour_viridis_c() +
          labs(x = "Longitude", y = "Latitude", color= "Daily\nSD(AOD)", 
               size = "Number of\nDays") +
          # ggrepel::geom_text_repel(data = aod_data[daysN>daysN0*1.5 & mediansd>median(aod_data$mediansd)],
          #                           aes(x = lon, y = lat, label = site),
          #                           size =2.5, inherit.aes = F) +
          theme_minimal() +
          ggtitle("SD by AERONET stations") +
          theme(legend.position ='right') 
  if(bySeason) g2 <- g2 + facet_wrap(~season)
  return(g2)
}
plotAODmap(aod2,daysN0 = 300)
p1 <- plotAODscatter(aod2)
p2 <- plotAODmap(aod2)
gridExtra::grid.arrange(grobs = list(p1, p2), ncol = 2)

```

### Calculating using past 3 years: 2015 - 2018

```{r}
# summary2: limiting the years! aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[year>=2015,.(lat, lon, dailyN, month, season, 
                            mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
                            year = year(date)), by = .(site, date)])
# 290 sites: aod2 [290 * 6]
aod1[, daysN := uniqueN(date), by = site]
aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site)])[yearmax>=2018]

# by season

aod3 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site, season)])[yearmax>=2018]

```

## make and save plot

```{r, fig.width=12, fig.height=5}
# print results with daysN > 500 
setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2)
p_aer <- plotAODmap(aod2, daysN0 = 300) + 
  ggtitle("Median Daily SD (AERONET AOD)", 
                   subtitle = "2015 - 2018")

# gridExtra::grid.arrange(grobs = list(p1, p_aer), ncol = 2)

D2 <- readRDS(here("Intermediate/DSCOVR_DailySD_AOD_bySite.rds"))
g_DS <- plotAODmap(D2, daysN0 = 5) + ggtitle("Median Daily SD (DSCOVR EPIC MAIAC AOD)", subtitle = "By AERONET Station, 2015 - 2018")

ggsave(p_aer, file = here("Figure","190731_DailySD_AER.png"),
      width = 6, height = 4)

ggsave(g_DS, file = here("Figure","190731_DailySD_DSCOVR.png"),
      width = 6, height = 4)

ggsave(grid.arrange(p_aer, g_DS, ncol = 2), file = here("Figure","190731_DailySD_DS_AERONET.png"),
      width = 12, height = 4)

```

By season

```{r, fig.width = 12, fig.height = 10}
plotAODscatter(aod3, bySeason = T)
plotAODmap(aod3, bySeason = T)
```

## Trend within day

### GSFC, Wallops (high sd) vs. "Sevilleta", "White_Sands_HELSTF" (low sd)

```{r, fig.width = 12, fig.height = 10}
# substract the daily mean -> 
aod_conus[, dailymean := mean(AOD_470nm), by = .(site, date)]
aod_conus[, aod_new := AOD_470nm - dailymean]
aod_conus2 <- aod_conus[site%in%c("Wallops", "GSFC",
                                  "Sevilleta", "White_Sands_HELSTF"),]
 # "MD_Science_Center", "SERC",

# by hour 
ggplot(dilute.data(aod_conus2,10), aes(x = hour , y = aod_new)) + 
  geom_point(size =1, alpha = 0.8) + 
  geom_smooth(se= F, method = "gam", formula = y~s(x, bs = "cs")) + 
  facet_wrap(~site, ncol = 2) +
  theme_classic()

# by date
ggplot(dilute.data(aod_conus2,10), aes(x = date , y = aod_new)) + 
  geom_point(size = 1, alpha = 0.8) + 
  geom_smooth(se= F, method = "gam", formula = y~s(x, bs = "cs")) + 
  facet_wrap(~site, ncol = 2) +
  theme_classic()

```

```{r }
# show daily trend: AOD vs date, all data
ggplot(data = aod1, aes(x = date, y = sd_aod, color = site)) +
  geom_point(size = 1) + 
  theme_classic() + theme(legend.position = "none") + 
  labs(y = "Daily SD(AOD) by station", x = "Date") + 
  ggtitle("Daily variation by station across time", subtitle = "(# Daily obs >20)")

```