### 2019-08-14 R/drake_plan.Rmd
new drake plan

---
title: "Drake Plan"
author: "Johnathan Rush, Yang Liu"
date: "08/14/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

```{r}
# See `drake_funcs.R` for functions used in drake plan #

# beginning of script where we will have a plan to:
# select Aeronet in CONUS
# select Aeronet in NEMIA
# calculate nearest idLSTpair0's. May need shortcut that uses buffer around stations to reduce possibility set.

# select from the daily data the subsets above (one of the two)
# calc columns as in old script
# as.numeric columns as necessary
# remove rows without AOD? probably not - they may have CWV
# ? which other columns to we want?
# restrict to 2000+ (parameter so that different subsets could be taken later)

# Libraries
suppressPackageStartupMessages({
  library(sf)
  library(magrittr)
  library(drake)
  library(data.table)
  library(fst)
  library(future)
  library(here)
  #install.packages("/home/rushj03/dev/nngeo", repos = NULL, type = "source")
  # library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
  #library(sp)
  #library(rgdal)
  #library(rnaturalearth) # REPLACE WITH CONUS FILE USED IN GRID XXX
  #library(rgeos)
  library(nngeo)
  #library(nngeo, lib.loc = "~/R/x86_64-pc-linux-gnu-library/3.6")
})

# Parameters ####

aer_stn_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/aeronet_locations_v3.txt"

aer_data_path = '~/qnap_geo/aeronet/All_Sites_Times_All_Points_AOD20_20180603.dat'
aer_data_path = '/scratch/temp/All_Sites_Times_All_Points_AOD20_20180603.dat' # don't have time to play with Ethernet all day
# the aer data with AOD470nm
aer_data_path_new = "~/Intermediate/AOD_data/190628_aod20_wPred_NEMIA_idLST.fst"

aer_files_path = "/data-coco/ECHO_PM/AeronetAODV3Level2/AOD/AOD20/ALL_POINTS/"
date_start = "2018-01-01" # do not include in sel_data_bytime if don't want to limit
date_end   = "2018-12-31" # do not include in sel_data_bytime if don't want to limit
geo_region = "NEMIA"
sats = c("terra", "aqua")
# MCD19 data: 
mcd19path = "/data-coco/mcd19/fst/nemia" # now there is only 2018 data 
# point to cache: 
ssd_cache = new_cache(path = "/scratch/cache/aeronet_drake")

# Drake Plans  ------------------------------------------------------------
# load functions for drake: 
source(here("R/drake_funcs.R"))
# new plan Yang Liu
nemia_plan <- drake_plan(
  # 1.  Aeroent ---------------------------------------------------------------------

  # AERONET station locations file: 
  aer_stns = fread(file = file_in(aer_stn_path), col.names = c("Site_Name", "lon", "lat", "elevm")),
  conus_buff = get_conus_buff(),
  nemia_buff = get_nemia_buff(),
  aerpts = get_aer_spatial(aer_stns),               # wgs84
  conus_aer = select_points(aerpts, conus_buff),    # AERONET sites in CONUS as points
  nemia_aer = select_points(conus_aer, nemia_buff), # AERONET sites in NEMIA as points

  # get nearest MODIS cells 
  refgrid = get_ref_grid(),
  candidate_refpts = points_in_buffer(nemia_aer, refgrid),
  aer_nearest = get_nearest_cell(nemia_aer, candidate_refpts), # Aeronet matched to grid, 174x4
  aer_nearest_NEMIA = remove_site_on_water(aer_nearest), # remove 2 sites

  # measurements. Editing values of date_start and date_end should make use of the cache
  # if range was previously calc'd
  aer_data = get_stn_data(aod_dir = file_in(aer_files_path)),  # 26171206x56, 11G
  aer_btw = target(sel_data_bytime(aer_data, date_start, date_end),
                   transform = map(date_start = !!date_start, date_end = !!date_end)),

  # selected measurements, transform by same date range if not automatic in name
  nemia = target(sel_data_bystation(aer_btw, nemia_aer),
                      transform = map(aer_btw)), # use map to map to aer_btw

  # 2. Interpolation -----------------------------------------------------------
  # so instead of running all the Aeronet data, only run what is used
  # it also join with the nearest grid id 
  wPred = target(interpolate_aod(nemia, aer_nearest_NEMIA),transform = map(nemia)),

  # 3. WPred Join MCD19 -----------------------------------------------------------
  # mcd is the large MCD19 dataset 
  mcd = target(read_lst(sat), transform = map(sat = c("terra", "aqua"))),
  # join is the rolling-joined dataset with aeronet
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred = wPred),
                transform = cross(mcd, wPred)),

  # 4. nemia_aer_m
  s = target(aer_in_mcd19(nemia_aer, mcd19 = join),
             transform = map(join)),
  # buffers 
  ref = target(ref_in_buffer(nemia_aer_m = s, refgrid),
             transform = map(s)),
  # sites
  p = target(get_site_list(nemia_aer_m = s), transform = map(s)),
  n = target(get_site_name(nemia_aer_m = s), transform = map(s)),

  # mapping
  # this way doesn't work:
  # sub = target(select_refgrid_subset(ref_list0 = ref, point0 = p, name0 = n),
  #              transform = map(ref, p, n))
  sub = target(purrr_map(ref = ref, p = p, n = n),
               transform = map(ref, p, n)),
  # calculate new vars 
  # for terra and aqua, using join, mcd, sub 
  new = target(aod_MODIS_newVars(aod_join_MODIS = join, 
                                 MODIS_all = mcd, refDT_sub = sub),
               transform = map(join, mcd, sub))
)
nemia_plan
#drake_plan_source(nemia_plan)
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
# vis_drake_graph(nemia_config, from = names(nemia_config$layout))
vis_drake_graph(nemia_config, ncol_legend = 0)

loadd(cache = ssd_cache)

message("To see one of the aer station with grid around it: ")
knitr::kable(sub_ref_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_p_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_n_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31[1:3,])

message("The colocated dataset with new vars")
rmarkdown::paged_table(new_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_mcd_aqua_sub_ref_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_p_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31_n_s_join_mcd_aqua_wPred_nemia_aer_btw_2018.01.01_2018.12.31[1:3,])
```

```{r, eval = F}
# outdated(nemia_config)
future::plan("multicore")
make(nemia_plan, cache = ssd_cache) # write update into the cache 
cached(cache = ssd_cache)
loadd(cache = ssd_cache) # load all object

# to see the updated result, run config again
nemia_config <- drake_config(nemia_plan, cache = ssd_cache) # show the dependency
vis_drake_graph(nemia_config)

# Join MCD19 --------------------------------------------------------------
aer_data_wPred <- readd("wPred_nemia_aer_btw_2018.01.01_2018.12.31", cache = ssd_cache) # load all object 
nemia_aer <- readd("nemia_aer", cache = ssd_cache)
refgrid <- readd("refgrid", cache = ssd_cache)

plan1 <- drake_plan(
  # 3. WPred Join MCD19 -----------------------------------------------------------
  mcd = target(read_lst(sat), transform = map(sat = c("terra", "aqua"))),
  join = target(rolling_join(mcd_sat = mcd, aer_data_wPred),
                transform = map(mcd))
)
config2 <- drake_config(plan1, cache = ssd_cache) # show the dependency
vis_drake_graph(config2)
time0 <- Sys.time()
make(plan1, cache = ssd_cache) # write update into the cache 
Sys.time() - time0
cached(cache = ssd_cache)
loadd(cache = ssd_cache) # load all object 
join_mcd_aqua = readd("join_mcd_aqua", cache = ssd_cache)

```