### 2019-08-18 AOD_CONUS_Report.Rmd
nemia drake

---
title: "AOD CONUS Modeling Results"
author: "Yang Liu"
date: "08/17/2019"
output: html_document
---

```{r}
library(drake)

```

# Terra

```{r}
readd(print1, cache = ssd_cache)
readd(plot1, cache = ssd_cache)
```

* Top features RFE

```{r}
# top features
loadd("cv_results_new_var_join_mcd_terra", cache = ssd_cache)
# rferesults = cv_results_new_var_join_mcd_terra
cv_results_new_var_join_mcd_terra$features_rank_rfe
```

```{r}

readd(plotSHAP1_2, cache = ssd_cache)

```

# Aqua

```{r}
readd(print1_2, cache = ssd_cache)
readd(plot1_2, cache = ssd_cache)

```

* Top features RFE

```{r}
# top features
loadd("cv_results_new_var_join_mcd_aqua", cache = ssd_cache)
cv_results_new_var_join_mcd_aqua$features_rank_rfe
```

```{r}
readd(plotSHAP1_2, cache = ssd_cache)

```