### 2019-08-22 AOD_CONUS_Report.Rmd
CONUS model using one-year data done

---
title: "AOD CONUS Modeling Results"
author: "Yang Liu"
date: "08/17/2019"
output: html_document
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

```{r}
library(drake) 
loadd(plotscatter, cache = ssd_cache)
n_vars = length(plotscatter)
fig_height = ceiling(n_vars/2)*3.5
```

# AOD Scatterplot

```{r, fig.width = 10, fig.height = 7}
loadd(plotAOD, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotAOD, ncol = 2)
```

# Terra

## RMSE

```{r}
readd(print1, cache = ssd_cache)
readd(plot1, cache = ssd_cache)
```

* Top features RFE

```{r}
# top features
loadd("cv_results_terra", cache = ssd_cache)
# rferesults = cv_results_new_var_join_mcd_terra
cv_results_terra$features_rank_rfe
```

## Scatterplot

```{r, fig.width = 10, fig.height = fig_height}
loadd(plotscatter, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotscatter, ncol = 2)
```

## SHAP

```{r}
readd(plotSHAP1, cache = ssd_cache)
```

```{r, fig.width = 10, fig.height = fig_height}
loadd(plotSHAP2, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotSHAP2, ncol = 2)
```

# Aqua

## RMSE

```{r}
readd(print1_2, cache = ssd_cache)
readd(plot1_2, cache = ssd_cache)
```

* Top features RFE

```{r}
# top features
loadd("cv_results_aqua", cache = ssd_cache)
cv_results_aqua$features_rank_rfe
```

## Scatterplot

```{r, fig.width = 10, fig.height = fig_height}
loadd(plotscatter_2, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotscatter_2, ncol = 2)

```

## SHAP

```{r}
readd(plotSHAP1_2, cache = ssd_cache)
```

```{r, fig.width = 10, fig.height = fig_height}
loadd(plotSHAP2_2, cache = ssd_cache)
gridExtra::grid.arrange(grobs = plotSHAP2_2, ncol = 2)

```