### 2024-02-05 writing/paper_aodc.qmd
Added AODC error envelopes.

---
title: aodc

bibliography: references.json
csl: https://raw.githubusercontent.com/citation-style-language/styles/795ad0c77258cb7e01f3413123b5b556b4cb6a98/dependent/journal-of-geophysical-research-atmospheres.csl

format:
  html: {embed-resources: true}

echo: FALSE
---

```{r}
source.files = "globals libraries util data modeling paper_functions"
for (filename in strsplit(source.files, " ")[[1]])
    source(sprintf("code/%s.R", filename))
suppressPackageStartupMessages(
    library(flextable))
IG = knitr::include_graphics
```

# Results

```{r}
d = copy(tar_read(cv)$mDT_wPred)
```

After collocation there were `r scales::comma(nrow(d))` matched AOD retrievals from `r uniqueN(d$site)` AERONET sites in CONUS. The quartiles of the absolute difference between the satellite time and the time of the matched AERONET observation were `r tqs = round(quantile(d[, abs(as.numeric(difftime(units = "secs", time.sat, time.ground)))], c(.25, .5, .75))); tqs["25%"]` s, `r tqs["50%"]` s, and `r tqs["75%"]` s.

@tbl-cv-overall includes the performance metrics for the full cross-validated training data set. `r sdn = \(x) sqrt(mean((x - mean(x))^2)); f = \(x) sprintf("(SD %.03f)", sdn(d[[x]])); "";` Here **raw** is … `r f("y.sat")`; **ground** is … `r f("y.ground")`; and **corrected** is … `r f("y.ground.pred")`. The ground values are correlated `r d[, sprintf("%+.03f", cor(y.ground, y.sat))]` with the raw values and `r d[, sprintf("%+.03f", cor(y.ground, y.ground.pred))]` with the corrected values.

```{r}
#| label: tbl-cv-overall
#| tbl-cap: Overall performance from cross-validation.
flextable(pretty.table.numbers(tar_read(cv.summary)[1, -"Year"]))
```

```{r}
#| label: tbl-cv-month
#| tbl-cap: Performance by month of year.
out = d[,
    keyby = .(Month = month(time.sat)),
    eval(performance.j)]
out[, Month := month.abb[Month]]
flextable(pretty.table.numbers(out))
```

```{r}
#| label: tbl-cv-hour
#| tbl-cap: Performance by hour of the day.
flextable(pretty.table.numbers(d[,
    keyby = .(Hour = seconds.since.midnight %/% (60L * 60L)),
    eval(performance.j)]))
```

```{r}
#| label: tbl-cv-high-aod
#| tbl-cap: Performance for higher versus lower AOD.
threshold = .6
flextable(pretty.table.numbers(d[,
    by = .("Ground value" = ifelse(y.ground > threshold,
        sprintf("> %.1f", threshold),
        sprintf("≤ %.1f", threshold))),
    eval(performance.j)]))
```

```{r}
#| label: tbl-cv-quality
#| tbl-cap: Performance by value of the quality variable `DQF`.
flextable(pretty.table.numbers(d[,
    keyby = .(DQF = ifelse(is.na(DQF), "missing", DQF)),
    eval(performance.j)]))
```

## Error envelopes

We use … The results are an envelope with `r eee = \(pred) {o = empirical.error.envelope(obs = d$y.ground, pred); stopifnot(o$convergence == 0); o$par}; show.eee = \(pred) {ps = eee(pred); sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; show.eee(d$y.sat)` for the raw satellite values, and an envelope with `r show.eee(d$y.ground.pred)` for our predictions. The mean widths of the envelopes across all observations are `r me = \(pred) {ps = eee(pred); sprintf("%.03f", 2*ps[1] + 2*ps[2]*mean(d$y.ground))}; me(d$y.sat)` for the raw values and `r me(d$y.ground.pred)` for our predictions.

```{r agreement_plot}
#| label: fig-agreement-plot
#| fig-cap: !expr caption
caption = with(list(d = agreement.plot.data(d)), sprintf(
    "Agreement plots for satellite-based and AERONET AOD. Straight lines show empirical error envelopes containing 2/3 of retrievals. Not shown: %s %s and %s %s",
    d[!in.region & comparison == "y.sat", .N],
    "points outside left panel",
    d[!in.region & comparison == "y.ground.pred", .N],
    "points outside right panel"))
IG(tar_read(agreement.plot.path))
```